---
title: "House prices - EDA and Modeling"
author: "stat17_hb"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    number_sections: true
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=T)
```

# Load libraries and data

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(plyr)
library(psych)
library(caret)
library(randomForest)
library(e1071)
library(xgboost)
library(neuralnet)
```


```{r, eval=FALSE}
# for kaggle kernel
list.files()
train <- read_csv("../input/train.csv")
test <- read_csv("../input/test.csv")
```

```{r, message=FALSE, warning=FALSE}
# for local env
list.files()
train <- read_csv("all/train.csv")
test <- read_csv("all/test.csv")
```


# Simple EDA before missing value imputation

## Glimpse data

```{r}
glimpse(train)
```

```{r}
glimpse(test)
```

+ Saleprice is target variable. It is only in the train data.


```{r}
# Combine train and test data for EDA
all <- dplyr::bind_rows(train, test)
```


## Classify categrical variables and continuous variables

```{r}
cat_var <- names(all)[which(sapply(all, is.character))]
numeric_var <- names(all)[which(sapply(all, is.numeric))]
numeric_var
```


```{r}
table(all$MSSubClass)
```

MSSubClass: Identifies the type of dwelling involved in the sale.

        20	1-STORY 1946 & NEWER ALL STYLES
        30	1-STORY 1945 & OLDER
        40	1-STORY W/FINISHED ATTIC ALL AGES
        45	1-1/2 STORY - UNFINISHED ALL AGES
        50	1-1/2 STORY FINISHED ALL AGES
        60	2-STORY 1946 & NEWER
        70	2-STORY 1945 & OLDER
        75	2-1/2 STORY ALL AGES
        80	SPLIT OR MULTI-LEVEL
        85	SPLIT FOYER
        90	DUPLEX - ALL STYLES AND AGES
       120	1-STORY PUD (Planned Unit Development) - 1946 & NEWER
       150	1-1/2 STORY PUD - ALL AGES
       160	2-STORY PUD - 1946 & NEWER
       180	PUD - MULTILEVEL - INCL SPLIT LEV/FOYER
       190	2 FAMILY CONVERSION - ALL STYLES AND AGES

+ 'MSSubClass' seems to be changed into categorical variable 


```{r}
cat_var <- c(cat_var, "MSSubClass")
numeric_var <- setdiff(numeric_var, intersect(cat_var, numeric_var))
cat_var
numeric_var
```

## EDA for continuous variables

### SalePrice(Target variable)

```{r, fig.width=8, fig.height=4}
par(mfrow=c(1,2))
hist(all$SalePrice, xlab="SalePrice", main="SalePrice")
hist(log(all$SalePrice), xlab="SalePrice", main="SalePrice")
```

+ SalePrice has skewed distribution.

+ Log transformation can change the right skewed distribution symmetrically.




### Histograms for continuous variables

```{r, fig.width=7, fig.height=7}
par(mfrow=c(3,3))
for (var in numeric_var[-c(1, 37)]){ # drop Id, SalePrice
  hist(unlist(all[,var]), main=var, xlab=var)
}
```


### Correlations with SalePrice

```{r, fig.width=10, fig.height=10}
cor.plot(all[numeric_var[-1]], xlas=2, diag=F, upper=F, cex=0.5)
```

+ OverallQual is highly correlated with SalePrice

```{r}
boxplot(all$SalePrice ~ all$OverallQual)
```




## EDA for categorical variables

### Barplots for categorical variables

```{r, fig.width=6, fig.height=6}
par(mfrow=c(3,3))
for (var in cat_var){
  barplot(table(all[, var]), main=var, las=2, cex.names=1)
}
```


# Missing value imputation

```{r}
colSums(is.na(all))
# colSums(sapply(all, is.na)) # same result
```


## Categorical varibles with missing values

```{r}
missing.cat_var <- cat_var[which(colSums(is.na(all[, cat_var])) != 0)]
```


```{r}
missing.cat.desc <- order(colSums(is.na(all[, cat_var]))[colSums(is.na(all[, cat_var]))!=0], 
                           decreasing = T)
colSums(is.na(all[, cat_var]))[colSums(is.na(all[, cat_var]))!=0][missing.cat.desc]
```

```{r}
## PoolQC
table(all$PoolQC)
```

PoolQC: Pool quality

       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       NA	No Pool

```{r}
all[is.na(all$PoolQC)==1 & all$PoolArea > 0, c("PoolQC", "PoolArea", "OverallQual")]
```

```{r}
all[all$PoolArea > 0, c("PoolQC", "PoolArea", "OverallQual")]
```

```{r}
all$PoolQC[is.na(all$PoolQC)==1 & all$PoolArea > 0] <- c("TA", "Gd", "TA")
```


```{r}
all[is.na(all$PoolQC), "PoolQC"] <- "No_Pool"
table(all$PoolQC)
```


```{r}
Qualities <- c('No_Pool' = 0, 'Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
all$PoolQC <- as.integer(revalue(all$PoolQC, Qualities))
table(all$PoolQC)
```


```{r}
## MiscFeature
table(all$MiscFeature)
```

MiscFeature: Miscellaneous feature not covered in other categories

       Elev	Elevator
       Gar2	2nd Garage (if not described in garage section)
       Othr	Other
       Shed	Shed (over 100 SF)
       TenC	Tennis Court
       NA	None
       
```{r}
all[is.na(all$MiscFeature), "MiscFeature"] <- "None"
table(all$MiscFeature)
```

```{r}
## Alley
table(all$Alley)
```

Alley: Type of alley access to property

       Grvl	Gravel
       Pave	Paved
       NA 	No alley access
       
       
```{r}
all[is.na(all$Alley), "Alley"] <- "No_alley"
table(all$Alley)
```

```{r}
## Fence
table(all$Fence)
```

Fence: Fence quality

       GdPrv	Good Privacy
       MnPrv	Minimum Privacy
       GdWo	Good Wood
       MnWw	Minimum Wood/Wire
       NA	No Fence
       
```{r}
all[is.na(all$Fence), "Fence"] <- "No_Fence"
table(all$Fence)
```

```{r}
## FireplaceQu
table(all$FireplaceQu)
```

FireplaceQu: Fireplace quality

       Ex	Excellent - Exceptional Masonry Fireplace
       Gd	Good - Masonry Fireplace in main level
       TA	Average - Prefabricated Fireplace in main living area or Masonry Fireplace in basement
       Fa	Fair - Prefabricated Fireplace in basement
       Po	Poor - Ben Franklin Stove
       NA	No Fireplace
       
       
```{r}
all[is.na(all$FireplaceQu), "FireplaceQu"] <- "No_Fireplace"
table(all$FireplaceQu)
```

```{r}
Qualities <- c('No_Fireplace' = 0, 'Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
all$FireplaceQu <- as.integer(revalue(all$FireplaceQu, Qualities))
table(all$FireplaceQu)
```

```{r}
## GarageFinish
table(all$GarageFinish)
```

GarageFinish: Interior finish of the garage

       Fin	Finished
       RFn	Rough Finished
       Unf	Unfinished
       NA	No Garage
       
```{r}
all[is.na(all$GarageFinish), "GarageFinish"] <- "No_Garage"
table(all$GarageFinish)
```

```{r}
## GarageQual
table(all$GarageQual)
```

GarageQual: Garage quality

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       NA	No Garage
       
```{r}
all[is.na(all$GarageQual), "GarageQual"] <- "No_Garage"
table(all$GarageQual)
```


```{r}
Qualities <- c('No_Garage' = 0, 'Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
all$GarageQual <- as.integer(revalue(all$GarageQual, Qualities))
table(all$GarageQual)
```


```{r}
## GarageCond
table(all$GarageCond)
```

GarageCond: Garage condition

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       NA	No Garage
       
       
```{r}
all[is.na(all$GarageCond), "GarageCond"] <- "No_Garage"
table(all$GarageCond)
```

```{r}
Qualities <- c('No_Garage' = 0, 'Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
all$GarageCond <- as.integer(revalue(all$GarageCond, Qualities))
table(all$GarageCond)
```



```{r}
## GarageType
table(all$GarageType)
```

GarageType: Garage location

       2Types	More than one type of garage
       Attchd	Attached to home
       Basment	Basement Garage
       BuiltIn	Built-In (Garage part of house - typically has room above garage)
       CarPort	Car Port
       Detchd	Detached from home
       NA	No Garage
       
```{r}
all[is.na(all$GarageType), "GarageType"] <- "No_Garage"
table(all$GarageType)
```

```{r}
## BsmtCond
table(all$BsmtCond)
```

BsmtCond: Evaluates the general condition of the basement

       Ex	Excellent
       Gd	Good
       TA	Typical - slight dampness allowed
       Fa	Fair - dampness or some cracking or settling
       Po	Poor - Severe cracking, settling, or wetness
       NA	No Basement
       
```{r}
all[is.na(all$BsmtCond), "BsmtCond"] <- "No_Basement"
table(all$BsmtCond)
```


```{r}
Qualities <- c('No_Basement' = 0, 'Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
all$BsmtCond <- as.integer(revalue(all$BsmtCond, Qualities))
table(all$BsmtCond)
```


```{r}
table(all$BsmtQual)
```

BsmtQual: Evaluates the height of the basement

       10 Ex	Excellent (100+ inches)
       7 Gd	Good (90-99 inches)
       5 TA	Typical (80-89 inches)
       3 Fa	Fair (70-79 inches)
       1 Po	Poor (<70 inches
       0 NA	No Basement

```{r}
all[is.na(all$BsmtQual), "BsmtQual"] <- "No_Basement"
table(all$BsmtQual) 
```


```{r}
Qualities <- c('No_Basement' = 0, 'Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
all$BsmtQual <- as.integer(revalue(all$BsmtQual, Qualities))
table(all$BsmtQual)
```


Bsmt- variables below here have the same NA type with BsmtCond

```{r}
## BsmtExposure
all[is.na(all$BsmtExposure), "BsmtExposure"] <- "No_Basement"
table(all$BsmtExposure)

## BsmtFinType2
all[is.na(all$BsmtFinType2), "BsmtFinType2"] <- "No_Basement"
table(all$BsmtFinType2) # NA : No Basement

## BsmtFinType1
all[is.na(all$BsmtFinType1), "BsmtFinType1"] <- "No_Basement"
table(all$BsmtFinType1) # NA : No Basement
```



```{r}
## MasVnrType
table(all$MasVnrType)
```

MasVnrType: Masonry veneer type

       BrkCmn	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       None	None
       Stone	Stone
       
```{r}
all[is.na(all$MasVnrType), "MasVnrType"] <- "None" # There are NAs that are not assigned to "None"
table(all$MasVnrType)
```

```{r}
table(all$MSZoning)
```

MSZoning: Identifies the general zoning classification of the sale.

       A	Agriculture
       C	Commercial
       FV	Floating Village Residential
       I	Industrial
       RH	Residential High Density
       RL	Residential Low Density
       RP	Residential Low Density Park
       RM	Residential Medium Density
       
```{r}
all[is.na(all$MSZoning), "MSZoning"] <- names(sort(table(all$MSZoning), decreasing = T))[1] # imputing mode
```

```{r}
table(all$Utilities)
```

Utilities: Type of utilities available

       AllPub	All public Utilities (E,G,W,& S)
       NoSewr	Electricity, Gas, and Water (Septic Tank)
       NoSeWa	Electricity and Gas Only
       ELO	Electricity only

+ almost all observations are in the "AllPub" category. So, I think it is not useful to predict house price. I dropped it.
       
```{r}
cat_var <- cat_var[cat_var != "Utilities"]
all <- all[,-which(names(all)=="Utilities")]
```

```{r}
table(all$Functional)
```

Functional: Home functionality (Assume typical unless deductions are warranted)

       Typ	Typical Functionality
       Min1	Minor Deductions 1
       Min2	Minor Deductions 2
       Mod	Moderate Deductions
       Maj1	Major Deductions 1
       Maj2	Major Deductions 2
       Sev	Severely Damaged
       Sal	Salvage only
       
```{r}
all[is.na(all$Functional), "Functional"] <- names(sort(table(all$Functional), decreasing = T))[1] # imputing mode
table(all$Functional)
```

```{r}
table(all$Exterior1st)
```

Exterior1st: Exterior covering on house

       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles

```{r}
all[is.na(all$Exterior1st), "Exterior1st"] <- names(sort(table(all$Exterior1st), decreasing = T))[1] # imputing mode
table(all$Exterior1st)
```

```{r}
table(all$Exterior2nd)
```

Exterior2nd: Exterior covering on house (if more than one material)

       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles

```{r}
all[is.na(all$Exterior2nd), "Exterior2nd"] <- names(sort(table(all$Exterior2nd), decreasing = T))[1] # imputing mode
table(all$Exterior2nd)
```

```{r}
table(all$Electrical)
```

Electrical: Electrical system

       SBrkr	Standard Circuit Breakers & Romex
       FuseA	Fuse Box over 60 AMP and all Romex wiring (Average)
       FuseF	60 AMP Fuse Box and mostly Romex wiring (Fair)
       FuseP	60 AMP Fuse Box and mostly knob & tube wiring (poor)
       Mix	Mixed

```{r}
all[is.na(all$Electrical), "Electrical"] <- names(sort(table(all$Electrical), decreasing = T))[1] # imputing mode
table(all$Electrical)
```

```{r}
table(all$KitchenQual)
```

KitchenQual: Kitchen quality

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor

```{r}
all[is.na(all$KitchenQual), "KitchenQual"] <- names(sort(table(all$KitchenQual), decreasing = T))[1] # imputing mode
table(all$KitchenQual)
```

```{r}
Qualities <- c('Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
all$KitchenQual <- as.integer(revalue(all$KitchenQual, Qualities))
table(all$KitchenQual)
```


```{r}
table(all$SaleType)
```

SaleType: Type of sale

       WD 	Warranty Deed - Conventional
       CWD	Warranty Deed - Cash
       VWD	Warranty Deed - VA Loan
       New	Home just constructed and sold
       COD	Court Officer Deed/Estate
       Con	Contract 15% Down payment regular terms
       ConLw	Contract Low Down payment and low interest
       ConLI	Contract Low Interest
       ConLD	Contract Low Down
       Oth	Other

```{r}
all[is.na(all$SaleType), "SaleType"] <- names(sort(table(all$SaleType), decreasing = T))[1] # imputing mode
table(all$SaleType)
```


```{r}
# missing values check tool
missing.cat.desc <- order(colSums(is.na(all[, cat_var]))[colSums(is.na(all[, cat_var]))!=0], 
                           decreasing = T)
colSums(is.na(all[, cat_var]))[colSums(is.na(all[, cat_var]))!=0][missing.cat.desc]
```


## Continuous varibles with missing values

```{r}
missing.numeric_var <- colSums(is.na(all[, numeric_var]))[colSums(is.na(all[, numeric_var]))!=0]
missing.numeric_var[order(missing.numeric_var, decreasing=T)]
```

```{r, fig.width=10, fig.height=5}
## LotFrontage: Linear feet of street connected to property
boxplot(all$LotFrontage~factor(all$Neighborhood), las=2, cex.axis=0.8)
```

```{r}
# impute with median by Neighborhood
LF_by_Nb <- aggregate(all$LotFrontage, by=list(factor(all$Neighborhood)), FUN=median, na.rm=T)

for (i in 1:nrow(LF_by_Nb)){
  all[is.na(all$LotFrontage) & all$Neighborhood == LF_by_Nb$Group.1[i], "LotFrontage"] <- LF_by_Nb$x[i]
}
```


```{r}
## MasVnrArea: Masonry veneer area in square feet
# need to check whether Masonry veneer exists
all$MasVnrType[is.na(all$MasVnrArea)] # None
all[is.na(all$MasVnrArea), "MasVnrArea"] <- 0 # impute with 0
```

```{r}
## BsmtFinSF1 : Type 1 finished square feet
# need to check whether Basement exists
all$BsmtFinType1[is.na(all$BsmtFinSF1)] # no basement
all$BsmtFinSF1[is.na(all$BsmtFinSF1)] <- 0

## BsmtFinSF2
all$BsmtFinType2[is.na(all$BsmtFinSF2)] # no basement
all$BsmtFinSF2[is.na(all$BsmtFinSF2)] <- 0

## BsmtUnfSF : Unfinished square feet of basement area
all$BsmtQual[is.na(all$BsmtUnfSF)] # no basement
all$BsmtUnfSF[is.na(all$BsmtUnfSF)] <- 0

## TotalBsmtSF : Total square feet of basement area
all$BsmtQual[is.na(all$TotalBsmtSF)] # no basement
all$TotalBsmtSF[is.na(all$TotalBsmtSF)] <- 0
```

```{r}
## GarageArea
all$GarageType[is.na(all$GarageArea)]
# imputing with median by GarageType
all$GarageArea[is.na(all$GarageArea)] <- median(all$GarageArea[all$GarageType=="Detchd"], na.rm=T)
```

```{r}
all$GarageType[is.na(all$GarageCars)]
# imputing with median by GarageType
all$GarageCars[is.na(all$GarageCars)] <- median(all$GarageCars[all$GarageType=="Detchd"], na.rm=T)
```

```{r}
table(all$GarageYrBlt)
```

+ 2207 seems to be typo.

```{r}
all$YearBuilt[which(all$GarageYrBlt==2207)]
all$YearRemodAdd[which(all$GarageYrBlt==2207)]
```

+ 2207 -> 2007

```{r}
all$GarageYrBlt[which(all$GarageYrBlt==2207)] <- 2007
```

```{r}
# need to check whether garage exists
all$GarageType[is.na(all$GarageYrBlt)]  # 2 Detchd, 157 NA(no garage)
```

+ I thought "GarageYrBlt" is not important variable to predict SalePrice. Because there are other variables containing year information. Also, it is hard to be used due to "No_Garage" elements. So, I dropped "GarageYrBlt". 

```{r}
all <- all[ ,-which(names(all) == "GarageYrBlt")]
numeric_var <- numeric_var[numeric_var!="GarageYrBlt"]
```


```{r}
all$BsmtCond[is.na(all$BsmtFullBath)]
```

BsmtFullBath: Basement full bathrooms

      0
      1
      2
      3

```{r}
all$BsmtFullBath[is.na(all$BsmtFullBath)] <- 0
```


```{r}
all$BsmtCond[is.na(all$BsmtHalfBath)]
```

BsmtHalfBath: Basement half bathrooms

      0
      1
      2

```{r}
all$BsmtHalfBath[is.na(all$BsmtHalfBath)] <- 0
```


```{r}
# missing values check tool
missing.numeric_var <- colSums(is.na(all[, numeric_var]))[colSums(is.na(all[, numeric_var]))!=0]
missing.numeric_var[order(missing.numeric_var, decreasing=T)]
```


```{r}
colSums(is.na(all))
```

+ Imputation is done.


# More EDA and Feature engineering

```{r}
time_var <- c("YearBuilt", "YearRemodAdd", "MoSold", "YrSold")
bstm_var <- c("BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF", "TotalBsmtSF")
bath_var <- c("BsmtFullBath", "BsmtHalfBath", "FullBath", "HalfBath")
garage_var <- c("GarageYrBlt", "GarageCars", "GarageArea")
porch_var <- c("OpenPorchSF", "EnclosedPorch", "3SsnPorch", "ScreenPorch")
etc <- c("KitchenAbvGr", "TotRmsAbvGrd", "Fireplaces", "MiscVal","WoodDeckSF")
numeric_var <- numeric_var[-1] # remove Id variable
```

```{r}
qualcond_var <- c("OverallQual", "OverallCond", "ExterQual","ExterCond",
                  "BsmtQual","BsmtCond","HeatingQC","KitchenQual",
                  "FireplaceQu","GarageQual","GarageCond","PoolQC", 
                  "BedroomAbvGr", "KitchenAbvGr", "TotRmsAbvGrd")
numeric_var <- numeric_var[-which(numeric_var %in% qualcond_var)]
cat_var <- cat_var[-which(cat_var %in% qualcond_var)]
```


## Transform character variables to factors

```{r}
all[cat_var] <- lapply(all[cat_var], factor)
sapply(all[, cat_var], class)
```

## Time variables : "HouseAge", "NewHouse"

    YearBuilt: Original construction date

    YearRemodAdd: Remodel date (same as construction date if no remodeling or additions)
    
    MoSold: Month Sold (MM)

    YrSold: Year Sold (YYYY)

+ HouseAge = YrSold - YearRemodAdd

```{r}
all$HouseAge <- all$YrSold - all$YearRemodAdd

unique(all$HouseAge)
all[all$HouseAge < 0, c("HouseAge", "YrSold", "YearRemodAdd", "YearBuilt")]
```

```{r}
all$HouseAge <- ifelse(all$YrSold - all$YearRemodAdd > 0, all$YrSold - all$YearRemodAdd, 0)
```


```{r}
fit.lm <- lm((all$SalePrice[which(!is.na(all$SalePrice))])~all$HouseAge[which(!is.na(all$SalePrice))])
summary(fit.lm)

fit.lm <- lm(log(all$SalePrice[which(!is.na(all$SalePrice))])~all$HouseAge[which(!is.na(all$SalePrice))])
summary(fit.lm)
```

```{r}
plot(all$HouseAge[which(!is.na(all$SalePrice))], log(all$SalePrice[which(!is.na(all$SalePrice))]),
     xlab="HouseAge", ylab="log(SalePrice)")
abline(fit.lm, col="red", lwd=2)
```

```{r}
numeric_var <- c(numeric_var, "HouseAge")
```


```{r}
all$IsRemodel <- factor(ifelse(all$YearBuilt == all$YearRemodAdd, 0, 1))
cat_var <- c(cat_var, "IsRemodel")
```


```{r}
all$Newhouse <- factor(ifelse(all$YrSold == all$YearBuilt, 1, 0))
cat_var <- c(cat_var, "Newhouse")
```


## GrLivArea

```{r}
fit.lm <- lm(SalePrice~GrLivArea, data=all[!is.na(all$SalePrice), ])
plot(all$SalePrice[!is.na(all$SalePrice)]~all$GrLivArea[!is.na(all$SalePrice)], xlab="GrLivArea", ylab="SalePrice")
abline(fit.lm, col="red", lwd=2, lty="dotted")
text(x=all$GrLivArea[!is.na(all$SalePrice)], y=all$SalePrice[!is.na(all$SalePrice)]+50000, labels = ifelse(all$GrLivArea[!is.na(all$SalePrice)] > 4500, rownames(all[!is.na(all$SalePrice),]), ""))
```

+ 524th, 1299th observations are outliers.

```{r}
all[c(524,1299), c("SalePrice", "GrLivArea", "OverallQual")]
```


## Square feet variables

> TotalBsmtSF: Total square feet of basement area

```{r}
cor(all$TotalBsmtSF, rowSums(all[c("BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF")]))
```

+ TotalBsmtSF covers "BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF".

> GrLivArea: Above grade (ground) living area square feet

```{r}
sf_var <- c(porch_var, "1stFlrSF", "2ndFlrSF", "LowQualFinSF")
cor(all$GrLivArea, rowSums(all[sf_var]))
```

```{r}
all$TotalSF <- all$TotalBsmtSF + all$GrLivArea
numeric_var <- c(numeric_var, "TotalSF")
```


```{r}
fit.lm <- lm(all$SalePrice[which(!is.na(all$SalePrice))]~all$TotalSF[which(!is.na(all$SalePrice))])
summary(fit.lm)
plot(all$TotalSF[which(!is.na(all$SalePrice))], all$SalePrice[which(!is.na(all$SalePrice))],
     xlab="TotalSF", ylab="SalePrice")
abline(fit.lm, col="red", lwd=2)
text(x=all$TotalSF[!is.na(all$SalePrice)], y=all$SalePrice[!is.na(all$SalePrice)]+50000, labels = ifelse(all$TotalSF[!is.na(all$SalePrice)] > 7500, rownames(all[!is.na(all$SalePrice),]), ""))
```



## Bath variables

```{r}
bath_var
all$TotalBath <- all$BsmtFullBath + 0.5*all$BsmtHalfBath + all$FullBath + 0.5*all$HalfBath
numeric_var <- c(numeric_var[-which(numeric_var %in% bath_var)], "TotalBath")
```

```{r}
fit.lm <- lm(all$SalePrice[which(!is.na(all$SalePrice))]~all$TotalBath[which(!is.na(all$SalePrice))])
summary(fit.lm)
plot(all$TotalBath[which(!is.na(all$SalePrice))], all$SalePrice[which(!is.na(all$SalePrice))],
     xlab="TotalSF", ylab="SalePrice")
abline(fit.lm, col="red", lwd=2)
```

## Porch variables

```{r}
porch_var
all$TotalPorchSF <- rowSums(all[porch_var])
numeric_var <- c(numeric_var, "TotalPorchSF")
```

```{r}
fit.lm <- lm(log(all$SalePrice[which(!is.na(all$SalePrice))])~all$TotalPorchSF[which(!is.na(all$SalePrice))])
summary(fit.lm)
plot(all$TotalPorchSF[which(!is.na(all$SalePrice))], log(all$SalePrice[which(!is.na(all$SalePrice))]),
     xlab="TotalPorchSF", ylab="log(SalePrice)")
abline(fit.lm, col="red", lwd=2)
```



```{r}
all <- all[-c(524, 1299),]
```



## Quality and Condition variables

```{r}
qualcond_df <- all[qualcond_var]
str(qualcond_df)
```

ExterQual: Evaluates the quality of the material on the exterior

       10 Ex	Excellent
       7 Gd	Good
       5 TA	Average/Typical
       3 Fa	Fair
       1 Po	Poor

```{r}
Qualities <- c('Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
qualcond_df$ExterQual <- as.integer(revalue(qualcond_df$ExterQual, Qualities))
table(qualcond_df$ExterQual)
```

ExterCond: Evaluates the present condition of the material on the exterior

       10 Ex	Excellent
       7 Gd	Good
       5 TA	Average/Typical
       3 Fa	Fair
       1 Po	Poor

```{r}
Qualities <- c('Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
qualcond_df$ExterCond <- as.integer(revalue(qualcond_df$ExterCond, Qualities))
table(qualcond_df$ExterCond)
```

HeatingQC: Heating quality and condition

       10 Ex	Excellent
       7 Gd	Good
       5 TA	Average/Typical
       3 Fa	Fair
       1 Po	Poor
       
```{r}
Qualities <- c('Po' = 1, 'Fa' = 3, 'TA' = 5, 'Gd' = 7, 'Ex' = 10)
qualcond_df$HeatingQC <- as.integer(revalue(qualcond_df$HeatingQC, Qualities))
table(qualcond_df$HeatingQC)
```      

```{r}
all$HasPool <- factor(ifelse(all$PoolArea > 0, 1, 0))
cat_var <- c(cat_var, "HasPool")
```

```{r}
boxplot(log(all$SalePrice)~all$HasPool, xlab="HasPool", ylab="log(SalePice)")
```

```{r}
t.test(log(all$SalePrice)~all$HasPool)
```

```{r}
med_SP_by_NB<- aggregate(SalePrice~Neighborhood, all, FUN=median)
med_SP_by_NB <- med_SP_by_NB[order(med_SP_by_NB$SalePrice), ]
barplot(height = med_SP_by_NB$SalePrice, names.arg = med_SP_by_NB$Neighborhood, las=2)
abline(h= median(all$SalePrice, na.rm=T), col="red", lty="dotted")
```

```{r}
quantile(all$SalePrice, na.rm=T)
```

```{r}
mean(all$SalePrice, na.rm=T)
```

```{r, fig.width=5, fig.height=5}
all$Neighbor_bin <- factor(ifelse(all$Neighborhood %in% as.character(med_SP_by_NB$Neighborhood[1:12]), 1,
                                  ifelse(all$Neighborhood %in% as.character(med_SP_by_NB$Neighborhood[13:22]), 2, 3)))

table(all$Neighbor_bin)
boxplot(all$SalePrice~all$Neighbor_bin, xlab="Neighorhood after binning", ylab="SalePrice")
summary(aov(SalePrice~Neighbor_bin, data=all))
cat_var <- c(cat_var, "Neighbor_bin")
```


```{r}
numeric_df <- all[numeric_var]
```


```{r, fig.width=7, fig.height=7}
par(mfrow=c(3,3))
for (var in numeric_var[-which(numeric_var=="SalePrice")]){
  hist(unlist(numeric_df[,var]), main=var, xlab=var)
}
```

```{r, fig.width=7, fig.height=7}
par(mfrow=c(3,3))
for (var in numeric_var){
  hist(log(unlist(numeric_df[,var])+1), main=var, xlab=var)
}
```


## Skewness adjustment and Scaling for continuous variables

```{r}
for(i in 1:ncol(numeric_df[-which(names(numeric_df)=="SalePrice")])){
        if (abs(skew(numeric_df[,i]))>0.8){
                numeric_df[,i] <- log(numeric_df[,i] +1)
        }
}

numeric_df[-which(names(numeric_df)=="SalePrice")] <- scale(numeric_df[-which(names(numeric_df)=="SalePrice")])

colSums(is.na(numeric_df))
```



## Dummy coding for categorical variables

```{r}
categ_df <- all[cat_var]
dummy_mat <- model.matrix(~.-1, categ_df)
dim(dummy_mat)

# remove variables that almost all values are 1s or 0s
sum(colSums(dummy_mat)/nrow(all) < 0.01)
sum(colSums(dummy_mat)/nrow(all) > 0.99)
idx <- which(colSums(dummy_mat)/nrow(all) >= 0.01 & colSums(dummy_mat)/nrow(all) <= 0.99)
dummy_df <- as.data.frame(dummy_mat[,idx])
```


## Combine numeric_df, dummy_df and qualcond_df + split to train and test

```{r}
combined <- cbind(numeric_df, dummy_df, qualcond_df)
X_train <- combined[!is.na(all$SalePrice), -which(names(combined)=="SalePrice")]
Y_train <- combined[!is.na(all$SalePrice), which(names(combined)=="SalePrice")]
X_test <- combined[is.na(all$SalePrice), -which(names(combined)=="SalePrice")]
```


# Variable selection

## Variable importance from Random Forest

```{r, fig.width=10, fig.height=10}
fit.rf <- randomForest(x=X_train, y=Y_train, importance=T)
varimp.rf <- importance(fit.rf)
varImpPlot(fit.rf)

IncMSE <- varimp.rf[,1]
top30 <- names(IncMSE[order(IncMSE, decreasing = T)][1:30])
```


# Modeling

```{r, eval=F}
fit.rf <- randomForest(x=X_train, y=Y_train)
pred.rf <- predict(fit.rf, newdata = X_test)
head(pred.rf)

submission_rf_log <- read.csv('all/sample_submission.csv')
submission_rf_log$SalePrice <- exp(pred.rf)
write.csv(submission_rf_log, file='submission_rf_log.csv', row.names = F)
```


```{r, eval=F}
fit.rf <- randomForest(x=X_train[,top30], y=Y_train)
pred.rf <- predict(fit.rf, newdata = X_test)
head(pred.rf)

submission_rf_log <- read.csv('all/sample_submission.csv')
submission_rf_log$SalePrice <- exp(pred.rf)
write.csv(submission_rf_log, file='submission_rf_top30.csv', row.names = F)
```

```{r, eval=F}
fit.svm <- svm(x=X_train, y=log(Y_train), kernel = "linear")
pred.svm <- predict(fit.svm, newdata = X_test)
head(exp(pred.svm))

submission_svm_log <- read.csv('all/sample_submission.csv')
submission_svm_log$SalePrice <- exp(pred.svm)
write.csv(submission_svm_log, file='submission_svm_notr.csv', row.names = F)
```


```{r, eval=F}
fit.svm <- svm(x=X_train, y=Y_train, kernel = "linear")
pred.svm <- predict(fit.svm, newdata = X_test)
head(exp(pred.svm))

submission_svm_log <- read.csv('all/sample_submission.csv')
submission_svm_log$SalePrice <- exp(pred.svm)
write.csv(submission_svm_log, file='submission_svm.csv', row.names = F)
```


```{r, eval=F}
fit.svm <- svm(x=X_train[,top30], y=Y_train, kernel = "linear")
pred.svm <- predict(fit.svm, newdata = X_test[,top30])
head(exp(pred.svm))

submission_svm_log <- read.csv('all/sample_submission.csv')
submission_svm_log$SalePrice <- exp(pred.svm)
write.csv(submission_svm_log, file='submission_svm_top30.csv', row.names = F)
```


```{r, eval=F}
fit.xgb <- xgboost(data=as.matrix(X_train), nrounds=700, label=Y_train, objective="reg:linear")
pred.xgb <- predict(fit.xgb, as.matrix(X_test))
head(pred.xgb)

submission_xgb_log <- read.csv('all/sample_submission.csv')
submission_xgb_log$SalePrice <- exp(pred.xgb)
write.csv(submission_xgb_log, file='submission_xgb_log.csv', row.names = F)
```



```{r, eval=F}
names(X_train) <- paste0("X", 1:length(names(X_train)))
train <- cbind(X_train, Y_train)
fit.nn <- neuralnet(as.formula(paste("Y_train ~", paste(names(X_train), collapse="+"))), data=train, hidden=c(10), rep=1, linear.output=T)
pred.nn <- compute(fit.nn, X_test)$net.result
head(exp(pred.nn))

submission_nn <- read.csv('all/sample_submission.csv')
submission_nn$SalePrice <- exp(pred.nn)
write.csv(submission_nn, file='submission_nn.csv', row.names = F)
```

